---
title: "Risk Tool Prototype"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    logo: images/one-food-logo.png
    theme:
      bg: "#ffffff"
      fg: "#2a398e" 
      primary: "#9fdbf5"
      base_font:
        google: Open Sans
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(readxl)
library(here)
library(data.table)
library(DT)
library(stringr)
library(patchwork)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

Introduction
================================================================================

Column
--------------------------------------------------------------------------------
#### One Food project

#### One Food risk tool

The One Food risk tool maintainer is Sarah Alewijnse (Cefas).
Please email sarah.alewijnse@cefas.gov.uk for debugging assistance.

How to
================================================================================

Column
--------------------------------------------------------------------------------

#### How to use the One Food risk tool

The One Food risk tool has two key components. 
First is the **Risk score template**, which allows users to create and download a custom
template for risk scores, based on the sectors, hazards, and impacts of their choice.
Second is the **Risk tool visualiser** which allows users to visualise those risk scores.

##### 1. Download the risk score template

Navigate to **Hazard and sector selection** under **Risk scores template**.
Select your required sector(s) and hazard(s) from the tick boxes.
Note that hazards are split into their categories: Biological, Chemical, Physical and Socio-Economic.
You can select as many hazards and sectors as you require.

Next, navigate to **Data template download** under **Risk scores template**. 
Here you can view your data template and download it as a .csv file.

##### 2. Fill in the risk scores

Using your preferred spreadsheet editing software, enter the risk scores for each hazard and sector, using the template you created.

##### 3. Upload the risk scores data

Navigate to **Data upload** under **Risk tool visualiser**. 
Upload the risk scores.

##### 4. Visualise the risk scores

Once you have uploaded the risk scores, you can visualise the results in **Risk visualiser** under **Risk tool visualiser**.
You can also view the risk scores data in the data tab.

Step 1 - Sector Selection {data-navmenu="Risk tool"}
================================================================================

Column {data-width=650}
--------------------------------------------------------------------------------

##### Step 1 of 4: Sector selection

Please input your sector(s).
Please separate multiple sectors with a comma.

```{r}
textInput(inputId = "input_sector",
                   label = "Sector(s):",
          placeholder = "Plese entor your sector(s)")

# Store upload in reactive value
template_creation <- reactiveValues(user = NULL)

# Load data
observe({
  # Ensure upload has been set
  req(input$input_sector)
  
  # Split string
  sector_split <- stringr::str_split_1(input$input_sector, ", ")
  
  # Assign to reactive value
  template_creation$template_sector <- sector_split
})
```

You have selected:

```{r}
# View
renderText({
  req(input$input_sector)
  input$input_sector
})
```

Step 2 - Hazard selection {data-navmenu="Risk tool"}
================================================================================

Column {data-width=650 .tabset}
--------------------------------------------------------------------------------

#### Step 2 of 4: Hazard selection

```{r}
checkboxGroupInput(inputId = "template_haz_bio",
                   label = "Biological hazards",
                   choices = c("Pathogens - Fungal"),
                   selected = c("Pathogens - Fungal"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_chem",
                   label = "Chemical hazards",
                   choices = c("Antimicrobials"),
                   selected = c("Antimicrobials"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_phys",
                   label = "Physical hazards",
                   choices = c("Water scarcity"),
                   selected = c("Water scarcity"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_soc",
                   label = "Socio-economic hazards",
                   choices = c("Food price inflation"),
                   selected = c("Food price inflation"))
```

Step 3 - Hazard Scoring {data-navmenu="Risk tool"}
================================================================================

Column {data-width=650 .tabset}
--------------------------------------------------------------------------------

#### View template

```{r}
# Base template ----------------------------------------------------------------
template_data <- data.frame(Category = NA,
                            Hazard = NA,
                            Impact = NA,
                            Sector = NA,
                            Severity = NA,
                            Probability = NA,
                            Longevity = NA)

# List of impacts/pillars ------------------------------------------------------
template_pillars <- reactive({data.frame(Pillar = c(rep("Animal/plant health", length(template_creation$template_sector)),
                                           rep("Environmental health", 3),
                                           rep("Human health", 3)),
                                Impact = c(template_creation$template_sector,
                                           "Biodiversity",
                                           "Pollution",
                                           "Carbon footprint",
                                           "Physical health",
                                           "Food availability",
                                           "Income"))
})

# Create template from inputs
template_data <- reactive({
  # Biological
  ifelse(length(input$template_haz_bio) > 0,
         biological_template <- data.frame(Category = rep("Biological", 
                                                          length(input$template_haz_bio) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_bio,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_bio)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_bio)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         biological_template <- data.frame(Category = NULL,
                                           Hazard = NULL,
                                           Pillar = NULL,
                                           Impact = NULL,
                                           Severity = NULL,
                                           Probability = NULL,
                                           Longevity = NULL))
  
  # Chemical
  ifelse(length(input$template_haz_chem) > 0,
         chemical_template <- data.frame(Category = rep("Chemical", 
                                                          length(input$template_haz_chem) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_chem,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_chem)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_chem)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         chemical_template <- data.frame(Category = NULL,
                                         Hazard = NULL,
                                         Pillar = NULL,
                                         Impact = NULL,
                                         Severity = NULL,
                                         Probability = NULL,
                                         Longevity = NULL))
  
  # Physical
  ifelse(length(input$template_haz_phys) > 0,
         physical_template <- data.frame(Category = rep("Physical", 
                                                          length(input$template_haz_phys) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_phys,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_phys)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_phys)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         physical_template <- data.frame(Category = NULL,
                                         Hazard = NULL,
                                         Pillar = NULL,
                                         Impact = NULL,
                                         Severity = NULL,
                                         Probability = NULL,
                                         Longevity = NULL))
  
  # Socio-economic
  ifelse(length(input$template_haz_soc) > 0,
         socio_economic_template <- data.frame(Category = rep("Socio-economic", 
                                                          length(input$template_haz_soc) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_soc,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_soc)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_soc)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         socio_economic_template <- data.frame(Category = NULL,
                                               Hazard = NULL,
                                               Pillar = NULL,
                                               Impact = NULL,
                                               Severity = NULL,
                                               Probability = NULL,
                                               Longevity = NULL))
  rbind(biological_template,
        chemical_template,
        physical_template,
        socio_economic_template)
})

# Create download handler
output$template <- downloadHandler(filename = function(){
  paste0(Sys.Date(), "_risk_score_template.csv")
}, 
content = function(file){
  write.csv(template_data(), file, row.names = F)
})

renderDataTable({
  template_data()
})
```



Column {data-width=200}
--------------------------------------------------------------------------------

#### Download template

```{r}
# Download button --------------------------------------------------------------
output$download_button <- renderUI({downloadButton(outputId = "template",
                                                   label = "Download .csv")})
uiOutput("download_button")
```

Column
--------------------------------------------------------------------------------
#### Data upload

```{r}
# Upload button
fileInput(inputId = "upload",
          label = "Upload hazard scores:",
          accept = ".csv")

# Store upload in reactive value
hazard_scores <- reactiveValues(user = NULL)

# Load data
observe({
  # Ensure upload has been set
  req(input$upload)
  
  # Read file
  file <- read.csv(file = input$upload$datapath)
  
  # Convert to data table
  file <- data.table(file)
  
  # Assign to reactive value
  hazard_scores$user <- file
})

```

#### Score generation

Choose whether the final hazard score will be generatived by additive or multiplicative methods.

```{r}
# Additive or multiplicative score generation
selectInput(inputId = "method",
            label = "Score generation method:",
            choices = c("Additive",
                        "Multiplicative"))
```

Step 4 - Risk visualisation {data-navmenu="Risk tool"}
================================================================================

Column {data-width=350 .sidebar}
-----------------------------------------------------------------------

```{r}
# First input selector
# View by sector or by hazard category
selectInput(inputId = "type",
            label = "Type",
            choices = c("Hazard overview",
                        "Hazard category"))

# Second input selectors -------------------------------------------------------

# Options for hazard category
output$category <- renderUI({
  if(input$type == "Hazard category")
    selectInput(inputId = "category",
                label = "Hazard category",
                choices = c("Biological",
                            "Chemical",
                            "Physical",
                            "Socio-economic"))
})

# Display by sector or impact
output$fill <- renderUI({
  if(input$type %in% c("Hazard overview", "Hazard category")){
    selectInput(inputId = "fill",
                label = "Fill by pillar or impact",
                choices = c("Pillar",
                            "Impact"))
  }
})

# Render UIs
uiOutput("category")
uiOutput("fill")
```

Column {data-width=650 .tabset}
-----------------------------------------------------------------------

### Plot

```{r}
# Edit synthetic data
updated_data <- reactive({
  if(input$method == "Additive"){
    hazard_scores$user$Score <- hazard_scores$user$Severity + hazard_scores$user$Probability + hazard_scores$user$Longevity
  }
  if(input$method == "Multiplicative"){
    hazard_scores$user$Score <- hazard_scores$user$Severity * hazard_scores$user$Probability * hazard_scores$user$Longevity
  }
  if(input$type == "Hazard category"){
    if(input$fill == "Pillar"){
      hazard_scores$user[Category %in% input$category][, .(Score = sum(Score)), by = c("Pillar", "Hazard")]
    } else if(input$fill == "Impact"){
      hazard_scores$user[Category %in% input$category][, .(Score = sum(Score)), by = c("Impact", "Hazard")]
    }
  } else if(input$type == "Hazard overview"){
    if(input$fill == "Pillar"){
      hazard_scores$user[, .(Score = sum(Score)), by = c("Category", "Pillar")]
    } else if(input$fill == "Impact"){
      hazard_scores$user[, .(Score = sum(Score)), by = c("Category", "Impact")]
    }
  }
})


# Colour palettes --------------------------------------------------------------

## Sector =====
# Okabe-Ito palette
sector_cols <- c("#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7",
                       "#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
                             "#44AA99", "#999933", "#882255", "#661100", "#6699CC")

## Hazard category =====

category_cols <- c("#CC6677",
                            "#44AA99",
                            "#6699CC",
                            "#AA4499")
                            
names(category_cols) <- c("Biological",
                          "Chemical",
                          "Physical",
                          "Socio-economic")

## Pillars ====

pillar_cols <- c("#6699CC",
                          "#117733",
                          "#882255")
                          
names(pillar_cols) <- c("Animal/plant health",
                        "Environmental health",
                        "Human health")

# Code for radar plot
coord_radar <- function (theta = "x", start = 0, direction = 1){
        theta <- match.arg(theta, c("x", "y"))
        r <- if (theta == "x") 
            "y"
        else "x"
        ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
            direction = sign(direction),
            is_linear = function(coord) TRUE)
        }

# Plots
of_plot <- reactive({
  if(input$type == "Hazard category" &&
     input$fill == "Pillar"){
    ggplot(data = updated_data()) +

      geom_bar(aes(x = str_wrap(Hazard, 5), y = Score, fill = Pillar,
                   col = Pillar), 
               stat = "identity",
               position = "dodge",
               alpha = 0.8) +
      scale_fill_manual(values = pillar_cols) +
      scale_colour_manual(values = pillar_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
  } else if(input$type == "Hazard category" &&
            input$fill == "Impact"){
    ggplot(data = updated_data()) +
      geom_bar(aes(x = str_wrap(Hazard, 5), y = Score, fill = Impact,
                   col = Impact), 
               stat = "identity",
               position = "dodge",
               alpha = 0.8) +
      scale_fill_manual(values = impact_cols) +
      scale_colour_manual(values = impact_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
  } else if(input$type == "Hazard overview" &&
            input$fill == "Pillar"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Pillar,
                   col = Pillar),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = pillar_cols) +
      scale_colour_manual(values = pillar_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Pillar, group = Pillar,
               fill = Pillar)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = pillar_cols) +
      scale_fill_manual(values = pillar_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar 
  } else if(input$type == "Hazard overview" &&
            input$fill == "Impact"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Impact,
                   col = Impact),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = impact_cols) +
      scale_colour_manual(values = impact_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15)) +
      guides(fill = guide_legend(nrow = 2))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Impact, group = Impact,
               fill = Impact)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = impact_cols) +
      scale_fill_manual(values = impact_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar
  }
})

# Plot
renderPlot({
  of_plot()
})
```


### Data

```{r}
renderDataTable({
  updated_data()
})
```
