---
title: "Risk Tool Prototype"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    logo: images/one-food-logo.png
    theme:
      bg: "#ffffff"
      fg: "#2a398e" 
      primary: "#9fdbf5"
      base_font:
        google: Open Sans
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(readxl)
library(here)
library(data.table)
library(DT)
library(stringr)
library(openxlsx)
library(patchwork)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

Introduction {data-navmenu="Introduction"}
================================================================================

Column
--------------------------------------------------------------------------------
#### One Food project

#### One Food risk tool

The One Food risk tool maintainer is Sarah Alewijnse (Cefas).
Please email sarah.alewijnse@cefas.gov.uk for debugging assistance.

How to {data-navmenu="Introduction"}
================================================================================

Column
--------------------------------------------------------------------------------

#### How to use the One Food risk tool

The One Food risk tool has two key components. 
First is the **Risk score template**, which allows users to create and download a custom
template for risk scores, based on the sectors, hazards, and impacts of their choice.
Second is the **Risk tool visualiser** which allows users to visualise those risk scores.

##### 1. Download the risk score template

Navigate to **Hazard and sector selection** under **Risk scores template**.
Select your required sector(s) and hazard(s) from the tick boxes.
Note that hazards are split into their categories: Biological, Chemical, Physical and Socio-Economic.
You can select as many hazards and sectors as you require.

Next, navigate to **Data template download** under **Risk scores template**. 
Here you can view your data template and download it as a .csv file.

##### 2. Fill in the risk scores

Using your preferred spreadsheet editing software, enter the risk scores for each hazard and sector, using the template you created.

##### 3. Upload the risk scores data

Navigate to **Data upload** under **Risk tool visualiser**. 
Upload the risk scores.

##### 4. Visualise the risk scores

Once you have uploaded the risk scores, you can visualise the results in **Risk visualiser** under **Risk tool visualiser**.
You can also view the risk scores data in the data tab.

1 - Sector selection
================================================================================

Column {data-width=650}
--------------------------------------------------------------------------------

#### Step 1 of 5: Sector selection

Please input your sector(s).
Please separate multiple sectors with a comma.

```{r}
textInput(inputId = "input_sector",
                   label = "Sector(s):",
          placeholder = "Please entor your sector(s)")

# Store upload in reactive value
template_creation <- reactiveValues(user = NULL)

# Load data
observe({
  # Ensure upload has been set
  req(input$input_sector)
  
  # Split string
  sector_split <- stringr::str_split_1(input$input_sector, ", ")
  
  # Assign to reactive value
  template_creation$template_sector <- sector_split
})
```

You have selected:

```{r}
# View
renderText({
  req(input$input_sector)
  input$input_sector
})
```

2 - Hazard selection
================================================================================

Column {data-width=650 .tabset}
--------------------------------------------------------------------------------

#### Step 2 of 5: Hazard selection

```{r}
checkboxGroupInput(inputId = "template_haz_bio",
                   label = "Biological hazards",
                   choices = c("Pathogens - Fungal"),
                   selected = c("Pathogens - Fungal"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_chem",
                   label = "Chemical hazards",
                   choices = c("Antimicrobials"),
                   selected = c("Antimicrobials"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_phys",
                   label = "Physical hazards",
                   choices = c("Water scarcity"),
                   selected = c("Water scarcity"))
```


```{r}
checkboxGroupInput(inputId = "template_haz_soc",
                   label = "Socio-economic hazards",
                   choices = c("Food price inflation"),
                   selected = c("Food price inflation"))
```

3 - Hazard scoring
================================================================================

#### Step 3 of 5: Hazard scoring

Check the generated table for hazard scoring.
When happy, download the template and complete the scores in Excel.

Column {data-width=650 .tabset}
--------------------------------------------------------------------------------

### View template

```{r}
# Base template ----------------------------------------------------------------
template_data <- data.frame(Category = NA,
                            Hazard = NA,
                            Impact = NA,
                            Sector = NA,
                            Severity = NA,
                            Probability = NA,
                            Longevity = NA)

# List of impacts/pillars ------------------------------------------------------
template_pillars <- reactive({data.frame(Pillar = c(rep("Animal/plant health", length(template_creation$template_sector)),
                                           rep("Environmental health", 3),
                                           rep("Human health", 3)),
                                Impact = c(template_creation$template_sector,
                                           "Biodiversity",
                                           "Pollution",
                                           "Carbon footprint",
                                           "Physical health",
                                           "Food availability",
                                           "Income"))
})

# Create template from inputs
template_data <- reactive({
  # Biological
  ifelse(length(input$template_haz_bio) > 0,
         biological_template <- data.frame(Category = rep("Biological", 
                                                          length(input$template_haz_bio) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_bio,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_bio)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_bio)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         biological_template <- data.frame(Category = NULL,
                                           Hazard = NULL,
                                           Pillar = NULL,
                                           Impact = NULL,
                                           Severity = NULL,
                                           Probability = NULL,
                                           Longevity = NULL))
  
  # Chemical
  ifelse(length(input$template_haz_chem) > 0,
         chemical_template <- data.frame(Category = rep("Chemical", 
                                                          length(input$template_haz_chem) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_chem,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_chem)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_chem)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         chemical_template <- data.frame(Category = NULL,
                                         Hazard = NULL,
                                         Pillar = NULL,
                                         Impact = NULL,
                                         Severity = NULL,
                                         Probability = NULL,
                                         Longevity = NULL))
  
  # Physical
  ifelse(length(input$template_haz_phys) > 0,
         physical_template <- data.frame(Category = rep("Physical", 
                                                          length(input$template_haz_phys) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_phys,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_phys)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_phys)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         physical_template <- data.frame(Category = NULL,
                                         Hazard = NULL,
                                         Pillar = NULL,
                                         Impact = NULL,
                                         Severity = NULL,
                                         Probability = NULL,
                                         Longevity = NULL))
  
  # Socio-economic
  ifelse(length(input$template_haz_soc) > 0,
         socio_economic_template <- data.frame(Category = rep("Socio-economic", 
                                                          length(input$template_haz_soc) *
                                                            length(template_pillars()$Pillar)),
                                           Hazard = sort(rep(input$template_haz_soc,
                                                        length(template_pillars()$Pillar))),
                                           Pillar = rep(template_pillars()$Pillar,
                                                        length(input$template_haz_soc)),
                                           Impact = rep(template_pillars()$Impact,
                                                        length(input$template_haz_soc)),
                                           Severity = NA,
                                           Probability = NA,
                                           Longevity = NA),
         socio_economic_template <- data.frame(Category = NULL,
                                               Hazard = NULL,
                                               Pillar = NULL,
                                               Impact = NULL,
                                               Severity = NULL,
                                               Probability = NULL,
                                               Longevity = NULL))
  template <- rbind(biological_template,
        chemical_template,
        physical_template,
        socio_economic_template) %>% data.table()
  template[order(Pillar)]
})

# Create download handler
output$template <- downloadHandler(filename = function(){
  paste0(Sys.Date(), "_risk_score_template.xlsx")
}, 
content = function(file){
  write.xlsx(template_data(), file)
})

renderDataTable({
  template_data()
})
```



Column {data-width=200}
--------------------------------------------------------------------------------

#### Download template

```{r}
# Download button --------------------------------------------------------------
output$download_button <- renderUI({downloadButton(outputId = "template",
                                                   label = "Download .xlsx")})
uiOutput("download_button")
```

4 - Score upload
================================================================================

#### Step 4 of 5: Score upload

Column {data-width = 650 .tabset}
--------------------------------------------------------------------------------
#### Data upload

```{r}
# Upload button
fileInput(inputId = "upload",
          label = "Upload hazard scores:",
          accept = ".xlsx")

# Store upload in reactive value
hazard_scores <- reactiveValues(user = NULL)

# Load data
observe({
  # Ensure upload has been set
  req(input$upload)
  
  # Read file
  file <- read.xlsx(xlsxFile = input$upload$datapath,
                   sheet = 1)
  
  # Convert to data table
  file <- data.table(file)
  
  # Assign to reactive value
  hazard_scores$user <- file
})

```

#### Score generation

```{r}
# Add or multiplicative score generation
selectInput(inputId = "method",
            label = "How do you want to calculate the overall hazard score?",
            choices = c("Add",
                        "Multiply"))
```

5 - Risk visualisation
================================================================================

Column {data-width=350 .sidebar}
-----------------------------------------------------------------------

```{r}
# First input selector
# View by sector or by hazard category
selectInput(inputId = "type",
            label = "Overview or pillar:",
            choices = c("Overview",
                        "Pillar"))

# Second input selectors -------------------------------------------------------

# Options for pillar
output$pillar <- renderUI({
  if(input$type == "Pillar")
    selectInput(inputId = "pillar",
                label = "Which pillar:",
                choices = c("Animal/plant health",
                            "Environmental health",
                            "Human health"))
})

# Render UIs
uiOutput("pillar")
```

Column {data-width=650 .tabset}
-----------------------------------------------------------------------

### Plot

```{r}
# Edit synthetic data
updated_data <- reactive({
  if(input$method == "Add"){
    hazard_scores$user$Score <- hazard_scores$user$Severity + hazard_scores$user$Probability + hazard_scores$user$Longevity
  }
  if(input$method == "Multiply"){
    hazard_scores$user$Score <- hazard_scores$user$Severity * hazard_scores$user$Probability * hazard_scores$user$Longevity
  }
  if(input$type == "Overview"){
    hazard_scores$user[, .(Score = sum(Score)), by = c("Category", "Pillar")]
  } else if(input$type == "Pillar"){
    if(input$pillar == "Animal/plant health"){
      hazard_scores$user[Pillar == "Animal/plant health"][, .(Score = sum(Score)), by = c("Category", "Impact")]
    } else if(input$pillar == "Environmental health"){
      hazard_scores$user[Pillar == "Environmental health"][, .(Score = sum(Score)), by = c("Category", "Impact")]
    } else if(input$pillar == "Human health"){
      hazard_scores$user[Pillar == "Human health"][, .(Score = sum(Score)), by = c("Category", "Impact")]
    }
  }
})

# Colour palettes --------------------------------------------------------------

## Hazard category =====
category_cols <- c("#E69F00",
                            "#56B4E9",
                            "#009E73",
                            "#D55E00")
                            
names(category_cols) <- c("Biological",
                          "Chemical",
                          "Physical",
                          "Socio-economic")

## Pillars ====
pillar_cols <- c("#6699CC",
                          "#117733",
                          "#882255")
                          
names(pillar_cols) <- c("Animal/plant health",
                        "Environmental health",
                        "Human health")

## Impacts ====

# Sector / animal/plant health
  # Okabe-Ito palette
sector_cols <- c("#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
                       
# Environmental health
env_cols <- c("#44AA99",
                       "#332288",
                       "#117733")
                       
names(env_cols) <- c("Biodiversity",
                     "Pollution",
                     "Carbon footprint")

# Human health
hum_cols <- c("#f28d9d",
                       "#AA4499",
                       "#661100")

# Code for radar plot
coord_radar <- function (theta = "x", start = 0, direction = 1){
        theta <- match.arg(theta, c("x", "y"))
        r <- if (theta == "x") 
            "y"
        else "x"
        ggproto("CordRadar", CoordPolar, theta = theta, r = r, start = start, 
            direction = sign(direction),
            is_linear = function(coord) TRUE)
        }

# Plots
of_plot <- reactive({
  # Overview plot
  if(input$type == "Overview"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Pillar,
                   col = Pillar),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = pillar_cols) +
      scale_colour_manual(values = pillar_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Pillar, group = Pillar,
               fill = Pillar)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = pillar_cols) +
      scale_fill_manual(values = pillar_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar 
    # Animal plant health
  } else if(input$type == "Pillar" &&
            input$pillar == "Animal/plant health"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Impact,
                   col = Impact),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = sector_cols) +
      scale_colour_manual(values = sector_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Impact, group = Impact,
               fill = Impact)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = sector_cols) +
      scale_fill_manual(values = sector_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar 
    # Environmental health
  } else if(input$type == "Pillar" &&
            input$pillar == "Environmental health"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Impact,
                   col = Impact),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = env_cols) +
      scale_colour_manual(values = env_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Impact, group = Impact,
               fill = Impact)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = env_cols) +
      scale_fill_manual(values = env_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar 
  } else if(input$type == "Pillar" &&
            input$pillar == "Human health"){
    bar <- ggplot(data = updated_data()) +
      geom_bar(aes(x = Category, y = Score, fill = Impact,
                   col = Impact),
               stat = "identity",
               position = "stack",
               alpha = 0.8) +
      scale_fill_manual(values = hum_cols) +
      scale_colour_manual(values = hum_cols) +
      xlab("\nCategory") +
      ylab("Hazard score\n") +
      theme_light() +
      theme(legend.position = "bottom",
            text = element_text(size = 15))
    spider <- ggplot(data = updated_data(),
           aes(x = Category, y = Score, col = Impact, group = Impact,
               fill = Impact)) +
      geom_polygon(alpha = 0, linewidth = 1) +
      geom_point(size = 5) +
      scale_colour_manual(values = hum_cols) +
      scale_fill_manual(values = hum_cols) +
      xlab("") +
      ylab("Hazard score\n") +
      coord_radar() +
      theme_light() +
      theme(legend.position = "none",
            text = element_text(size = 15))
    spider + bar 
  }
})

# Plot
renderPlot({
  of_plot()
})
```


### Data

```{r}
renderDataTable({
  updated_data()
})
```
